CF Registry Analysis

### 2023-12-08 Christine Chan/CF Registry/cf_registry_analysis.qmd
CF registry interval censored data ready

```{r}
# Import data and codebook
annual <- read.csv("./Data_Raw/CFF19_Annualized_Del1.csv",
  na.strings = c("", "N/A")
)
data_dict <- readxl::read_excel("./Data_Raw/Copy of Codes for CFFPR_2018.xlsx")
data_dict$var <- paste(data_dict$field_name, data_dict$code_value, sep = "_")
# Remove those with T1D or T2D
annual <- annual %>% arrange(eDWID, ReviewYear)
annual <- annual[!annual$cfrd_status_annual %in% c(4, 5), ]
annual$cfrd_status_annual <- factor(annual$cfrd_status_annual,
  levels = 1:3,
  labels = c("CFRD-", "CFRD-", "CFRD+")
)
# Get mutations
demogs <- read.csv("./Data_Raw/CFF19_DemogCFDiag_Del1.csv")
demogs <- demogs %>% select(eDWID, Gender, MutClass)
annual <- left_join(annual, demogs, by = join_by(eDWID))
# Variable formatting
# If A_supplement_feeding is 0, then A_feeding2 is also 0 because they are not
# on any supplemental feeding
# Include all kinds of tube feeds
annual$A_feeding2[annual$A_supplement_feeding == 0] <- 0
annual$A_feeding3[annual$A_supplement_feeding == 0] <- 0
annual$A_feeding4[annual$A_supplement_feeding == 0] <- 0
tube_fields <- c("A_feeding2", "A_feeding3", "A_feeding4")
annual$tube_feeding <-
  apply(annual[, tube_fields], 1, function(r) {
    if (all(is.na(r))) {
      return(NA)
    } else {
      return(max(as.numeric(r), na.rm = T))
    }
  })
# Liver disease
liver_fields <- c("A_hepatobiliary1_3", paste0("A_hepatobiliary2_", 1:4))
annual$liver_disease <-
  apply(annual[, liver_fields], 1, function(r) {
    if (all(is.na(r))) {
      return(NA)
    } else {
      return(max(as.numeric(r), na.rm = T))
    }
  })
# Calculate percentiles
annual$bmi <- annual$A_weight / ((annual$A_height / 100)^2)
annual$age_perc <- ifelse(annual$Age_YrEnd < 20, annual$Age_YrEnd, 20)
annual$height_perc <- sds(annual$A_height,
  age = annual$age_perc,
  sex = annual$Gender, male = "M", female = "F",
  item = "height2_20", ref = cdc.ref, type = "perc"
) * 100
annual$bmi_perc <- sds(annual$bmi,
  age = annual$age_perc,
  sex = annual$Gender, male = "M", female = "F",
  item = "bmi", ref = cdc.ref, type = "perc"
) * 100
# Format columns
annual$MutClass <- factor(annual$MutClass,
  levels = c("1-3", "4-5", "Oth", "N/A"),
  labels = c("1-3", "4-5", "Oth", "Unknown")
)
annual$NumPulmExacerbation <- cut(annual$NumPulmExacerbation,
  labels = c("0", "1", "2", ">=3"),
  breaks = c(0, 1, 2, 3, Inf), right = F
)
annual$age_group <- cut(annual$Age_YrEnd,
  breaks = c(-Inf, 10, 18, 30, Inf),
  right = F
)
annual$age_group <- relevel(annual$age_group, ref = "[10,18)")
# Bugs
bugs <- c(
  "A_pseudomonasaeruginosa", "A_fungalyeast1",
  "A_pulmonarycomplications1", "A_mycocultureresults3"
)
# Yes/No variables
yn_vars <- c(
  "A_IsOnEnzymes", "A_corticosteroids1", "A_feeding2",
  "A_Vx770", "A_VX809comb", bugs, "tube_feeding", "liver_disease"
)
annual[, yn_vars] <- lapply(annual[, yn_vars], factor,
  levels = 0:1,
  labels = c("No", "Yes")
)
annual$A_IsOnEnzymes <- relevel(annual$A_IsOnEnzymes, ref = "Yes")
# Approximate year of birth
by <- annual %>%
  mutate(birth_year = floor(ReviewYear - Age_YrEnd)) %>%
  group_by(eDWID) %>%
  count(birth_year) %>%
  group_by(eDWID) %>%
  slice_min(birth_year) %>%
  select(-n)
annual <- left_join(annual, by, by = join_by(eDWID))
```

@@

```{r}
# List of covariates
vars <- c(
  "age_group", "Gender", "bmi_perc", "MutClass", "A_IsOnEnzymes",
  "A_FEV1_pct_predicted", "tube_feeding", "NumPulmExacerbation",
  "A_pseudomonasaeruginosa", "A_fungalyeast1", "A_mycocultureresults3",
  "A_pulmonarycomplications1", "A_corticosteroids1", "cfrd_status_annual",
  "A_Vx770", "A_VX809comb", "liver_disease"
)
labels <- list(
  cfrd_status_annual = "CFRD Status", MutClass = "Mutation class",
  Gender = "Sex", A_IsOnEnzymes = "Pancreatic enzyme use",
  A_FEV1_pct_predicted = "FEV1 percent predicted",
  tube_feeding = "Enteric feedings", liver_disease = "Liver disease",
  A_corticosteroids1 = "Use of corticosteroids",
  NumPulmExacerbation = "Admissions for PEx", age_group = "Age group",
  height_perc = "Height percentile", bmi_perc = "BMI percentile",
  A_Vx770 = "Use of CFTR modulator Vx770",
  A_VX809comb = "Use of CFTR modulator Vx809 combination",
  A_pseudomonasaeruginosa = "Pseudomonas aeruginosa",
  A_fungalyeast1 = "Aspergillus (any species)",
  A_pulmonarycomplications1 = "ABPA",
  A_mycocultureresults3 = "Microorganisms in myco culture"
)
# For now, let's just assume that everyone's first CFRD+ diagnosis is correct
s <- annual %>%
  filter(!is.na(cfrd_status_annual), Age_YrEnd != 0) %>%
  select(
    eDWID, ReviewYear, birth_year, Age_YrEnd,
    all_of(vars), cfrd_status_annual
  ) %>%
  rename(tstop = ReviewYear) %>%
  group_by(eDWID) %>%
  mutate(
    tstart = lag(tstop, default = birth_year[1]),
    age_at_entry = Age_YrEnd[1]
  ) %>%
  select(eDWID, birth_year, tstart, tstop, cfrd_status_annual, age_at_entry, everything())
write.csv(s, file = "~/survival_data.csv", row.names = F, na = "")
```

```{r}
cox_res <- coxph(
  Surv(tstart, tstop, cfrd_status_annual == "CFRD+") ~
    age_group + A_FEV1_pct_predicted,
  data = s
)
ggsurvplot(survfit(cox_res),
  data = s,
  risk.table = T, cumevents = T, fun = "event", xlim = c(2003, 2018)
)
```

>>>>

```{r}
# Import data and codebook
annual <- read.csv("./Data_Raw/CFF19_Annualized_Del1.csv",
  na.strings = c("", "N/A")
)
data_dict <- readxl::read_excel("./Data_Raw/Copy of Codes for CFFPR_2018.xlsx")
data_dict$var <- paste(data_dict$field_name, data_dict$code_value, sep = "_")
# Remove those with T1D or T2D
annual <- annual %>% arrange(eDWID, ReviewYear)
annual <- annual[!annual$cfrd_status_annual %in% c(4, 5), ]
annual$cfrd_status_annual <- factor(annual$cfrd_status_annual,
  levels = 1:3,
  labels = c("CFRD-", "CFRD-", "CFRD+")
)
# Get mutations
demogs <- read.csv("./Data_Raw/CFF19_DemogCFDiag_Del1.csv")
demogs <- demogs %>% select(eDWID, Gender, MutClass)
annual <- left_join(annual, demogs, by = join_by(eDWID))
# Variable formatting
# If A_supplement_feeding is 0, then A_feeding2 is also 0 because they are not
# on any supplemental feeding
# Include all kinds of tube feeds
annual$A_feeding2[annual$A_supplement_feeding == 0] <- 0
annual$A_feeding3[annual$A_supplement_feeding == 0] <- 0
annual$A_feeding4[annual$A_supplement_feeding == 0] <- 0
tube_fields <- c("A_feeding2", "A_feeding3", "A_feeding4")
annual$tube_feeding <-
  apply(annual[, tube_fields], 1, function(r) {
    if (all(is.na(r))) {
      return(NA)
    } else {
      return(max(as.numeric(r), na.rm = T))
    }
  })
# Liver disease
liver_fields <- c("A_hepatobiliary1_3", paste0("A_hepatobiliary2_", 1:4))
annual$liver_disease <-
  apply(annual[, liver_fields], 1, function(r) {
    if (all(is.na(r))) {
      return(NA)
    } else {
      return(max(as.numeric(r), na.rm = T))
    }
  })
# Calculate percentiles
annual$bmi <- annual$A_weight / ((annual$A_height / 100)^2)
annual$age_perc <- ifelse(annual$Age_YrEnd < 20, annual$Age_YrEnd, 20)
annual$height_perc <- sds(annual$A_height,
  age = annual$age_perc,
  sex = annual$Gender, male = "M", female = "F",
  item = "height2_20", ref = cdc.ref, type = "perc"
) * 100
annual$bmi_perc <- sds(annual$bmi,
  age = annual$age_perc,
  sex = annual$Gender, male = "M", female = "F",
  item = "bmi", ref = cdc.ref, type = "perc"
) * 100
# Format columns
annual$MutClass <- factor(annual$MutClass,
  levels = c("1-3", "4-5", "Oth", "N/A"),
  labels = c("1-3", "4-5", "Oth", "Unknown")
)
annual$NumPulmExacerbation <- cut(annual$NumPulmExacerbation,
  labels = c("0", "1", "2", ">=3"),
  breaks = c(0, 1, 2, 3, Inf), right = F
)
annual$age_group <- cut(annual$Age_YrEnd,
  breaks = c(-Inf, 10, 18, 30, Inf),
  right = F
)
annual$age_group <- relevel(annual$age_group, ref = "[10,18)")
# Bugs
bugs <- c(
  "A_pseudomonasaeruginosa", "A_fungalyeast1",
  "A_pulmonarycomplications1", "A_mycocultureresults3"
)
# Yes/No variables
yn_vars <- c(
  "A_IsOnEnzymes", "A_corticosteroids1", "A_feeding2",
  "A_Vx770", "A_VX809comb", bugs, "tube_feeding", "liver_disease"
)
annual[, yn_vars] <- lapply(annual[, yn_vars], factor,
  levels = 0:1,
  labels = c("No", "Yes")
)
annual$A_IsOnEnzymes <- relevel(annual$A_IsOnEnzymes, ref = "Yes")
# Approximate year of birth
by <- annual %>%
  mutate(birth_year = floor(ReviewYear - Age_YrEnd)) %>%
  group_by(eDWID) %>%
  count(birth_year) %>%
  group_by(eDWID) %>%
  slice_min(birth_year) %>%
  select(-n)
annual <- left_join(annual, by, by = join_by(eDWID))
# Get first year of CFRD
first_cfrd <- annual %>%
  filter(cfrd_status_annual == "CFRD+") %>%
  group_by(eDWID) %>%
  summarise(first_cfrd = min(ReviewYear))
annual <- left_join(annual, first_cfrd, by = join_by(eDWID))
```

@@

```{r}
# List of covariates
vars <- c(
  "age_group", "Gender", "bmi_perc", "MutClass", "A_IsOnEnzymes",
  "A_FEV1_pct_predicted", "tube_feeding", "NumPulmExacerbation",
  "A_pseudomonasaeruginosa", "A_fungalyeast1", "A_mycocultureresults3",
  "A_pulmonarycomplications1", "A_corticosteroids1", "cfrd_status_annual",
  "A_Vx770", "A_VX809comb", "liver_disease"
)
labels <- list(
  cfrd_status_annual = "CFRD Status", MutClass = "Mutation class",
  Gender = "Sex", A_IsOnEnzymes = "Pancreatic enzyme use",
  A_FEV1_pct_predicted = "FEV1 percent predicted",
  tube_feeding = "Enteric feedings", liver_disease = "Liver disease",
  A_corticosteroids1 = "Use of corticosteroids",
  NumPulmExacerbation = "Admissions for PEx", age_group = "Age group",
  height_perc = "Height percentile", bmi_perc = "BMI percentile",
  A_Vx770 = "Use of CFTR modulator Vx770",
  A_VX809comb = "Use of CFTR modulator Vx809 combination",
  A_pseudomonasaeruginosa = "Pseudomonas aeruginosa",
  A_fungalyeast1 = "Aspergillus (any species)",
  A_pulmonarycomplications1 = "ABPA",
  A_mycocultureresults3 = "Microorganisms in myco culture"
)
# For now, let's just assume that everyone's first CFRD+ diagnosis is correct
s <- annual %>%
  filter(
    !is.na(cfrd_status_annual),
    Age_YrEnd != 0,
    is.na(first_cfrd) | ReviewYear <= first_cfrd
  ) %>%
  select(
    eDWID, ReviewYear, birth_year, Age_YrEnd,
    all_of(vars), cfrd_status_annual
  ) %>%
  rename(tstop = ReviewYear) %>%
  group_by(eDWID) %>%
  mutate(
    tstart = lag(tstop, default = birth_year[1]),
    age_at_entry = Age_YrEnd[1]
  ) %>%
  select(eDWID, birth_year, tstart, tstop, cfrd_status_annual, age_at_entry, everything())
```

```{r}
cox_res <- coxph(
  Surv(tstart, tstop, cfrd_status_annual == "CFRD+") ~
    age_group + A_FEV1_pct_predicted,
  data = s, cluster = eDWID
)
ggsurvplot(survfit(cox_res),
  data = s,
  risk.table = T, cumevents = T, fun = "event", xlim = c(2003, 2018)
)
surv_fit = survfit(Surv(tstart, tstop, cfrd_status_annual == "CFRD+") ~ age_group, 
                   data = s)
ggsurvplot(surv_fit,
  risk.table = T, cumevents = T, fun = "event", xlim = c(2003, 2018)
)
```
